---
title: "lab3: Bionomial Logistic Regression"
author: "Nick McManus"
date: "2023-01-26"
output: html_document
---

```{r setup, echo = T, warning = F, message = F}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)
```


## Pseudocode
- Examine our data (plots, tables, summary stats)
- Identify a question
- wrangle data if necessary
- identify some candidate models
- select among candidate models using AIC/BIC
- select among candidate models using K-fold cross validation
- select among candidate models using area under Receiver Operating Charatersitic Curve


```{r}
GGally::ggpairs(penguins %>% select(species, bill_length_mm:sex),
                aes(color = species))
```


```{r}
class(penguins$species)
levels(penguins$species)

adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>% 
  #drops any factors in variable 'species' no longer around
  mutate(species = fct_drop(species)) %>% 
  select(-year) %>% 
  #drops any obs with NA
  drop_na()

#still includes Gentoo w/o mutate step
#levels(adelie_chinstrap$species)
```


## Trends across variables

```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~species)

ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm))+
  geom_point(aes(color = sex, shape = island))+
  facet_wrap(~species)

## bill length seems to separate species better than flipper length
## (better association with species type)
```



## Binary logistic regression

```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex

ad_chin_blr1 <- glm(formula = f1, data = adelie_chinstrap, family = 'binomial')

ad_chin_blr1
summary(ad_chin_blr1)
# body mass and sex don't seem to be sig factors in estimating species

#puts outputs into df for easier viewing
blr1_tidy <- tidy(ad_chin_blr1)
```


```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```


```{r}
blr1_fitted <- ad_chin_blr1 %>% 
  broom::augment(type.predict = 'response')

ggplot(data = blr1_fitted, aes(x=flipper_length_mm, y = .fitted)) +
  geom_point(aes(color = sex, shape = species)) +
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = 'Flipper length (mm)',
       y = 'Probability of outcome(Chinstrap)')
```




## Predictions for new values with 'predict()'

```{r}
ex1 <- predict(ad_chin_blr1, 
               data.frame(sex = 'female',
                          body_mass_g = 3410,
                          flipper_length_mm = 192),
               type = 'response')

ex1
#shows a 40% chance that this is a chinstrap penguin

new_df <- data.frame(
  sex = c('male', 'female', 'female'),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)

ex2 <- predict(ad_chin_blr1, new_df, type = 'response')
#shows the likelihood for each of the obs in this df
ex2
```

## Create a new binary logistic model

```{r}
f2 <- species ~ bill_length_mm + body_mass_g

ad_chin_blr2 <- glm(formula = f2, data = adelie_chinstrap, family = 'binomial')
ad_chin_blr2
summary(ad_chin_blr2)
blr2_tidy <- tidy(ad_chin_blr2)
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
```



## Model selection

```{r}
#compare the AICs of the two models
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2))
#compare the BICs
AICcmodavg::bictab(list(ad_chin_blr1, ad_chin_blr2))
```

10-fold cross validation
```{r}
#arbitrar, but same seed between diff ppl should get same outputs
set.seed(123)

n_folds <- 10
fold_vec <- rep(1:n_folds, length.out = nrow(adelie_chinstrap))

ad_chin_kfold <- adelie_chinstrap %>% 
  mutate(fold = sample(fold_vec, size = n(), replace = FALSE))

#look at last lab for info on how to write for-loop to do this
```

Instead of for loop, going to look at purrr package
`purrr::map()`

```{r}
x_vec <- 1:10
thing <- purrr::map(.x = x_vec, ### a sequence (vec, list),
                    .f = sqrt)  ### a function
thing



my_funct <- function(x, y, z) {
  return((x - y)^z)
}

thing2 <- purrr::map(.x = x_vec,
                    .f = my_funct,
                    y=2, z = 3)
thing2
```

```{r}
pred_acc <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  
  return(mean(accurate, na.rm = TRUE))
}

calc_fold <- function(i, fold_df, f) {
  kfold_test <- fold_df %>% 
    filter(fold == i)
  kfold_train <- fold_df %>% 
    filter(fold != i)
  
  kfold_blr <- glm(f, data = kfold_train, family = 'binomial')
  
  kfold_pred <- kfold_test %>% 
    mutate(blr = predict(kfold_blr, kfold_test, type = 'response'),
           pred = ifelse(blr > 0.5, 'Chinstrap', 'Adelie')) #threshold of 50% for prediction
  
  kfold_accuracy <- kfold_pred %>% 
    summarize(blr_acc = pred_acc(species, pred))
  
  return(kfold_accuracy)
}
```


```{r}
results1_purrr_df <- purrr::map(.x = 1:n_folds,
                                .f = calc_fold,
                                fold_df = ad_chin_kfold,
                                f = f1) %>% 
  bind_rows() %>% 
  mutate(mdl = 'f1')

results2_purrr_df <- purrr::map(.x = 1:n_folds,
                                .f = calc_fold,
                                fold_df = ad_chin_kfold,
                                f = f2) %>% 
  bind_rows() %>% 
  mutate(mdl = 'f2')


results_purrr_df <- bind_rows(results1_purrr_df, results2_purrr_df) %>% 
  group_by(mdl) %>% 
  summarize(mean_acc = mean(blr_acc))
```


# tidymodels version

```{r}
### define model type
blr_model <- logistic_reg() %>% 
  set_engine('glm')

### set a basic regression
blr_tidyfit_f1 <- blr_model %>% 
  fit(f1, data = adelie_chinstrap)
blr_tidyfit_f2 <- blr_model %>% 
  fit(f2, data = adelie_chinstrap)

blr_tidyfit_f1

blr_tidyfit_f1 %>% 
  tidy()
blr_tidyfit_f2 %>% 
  glance()
```


# Tidy kfold cross validation

```{r}
set.seed(345)

tidy_folds <- vfold_cv(adelie_chinstrap, v = 10)

## use workflow to bundle a model and a formula
blr_tidy_wf1 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f1)

blr_tidy_cv_f1 <- blr_tidy_wf1 %>% 
  fit_resamples(tidy_folds)

collect_metrics(blr_tidy_cv_f1)

blr_tidy_wf2 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f2)

blr_tidy_cv_f2 <- blr_tidy_wf2 %>% 
  fit_resamples(tidy_folds)

collect_metrics(blr_tidy_cv_f2)
```


## Area under the curve

```{r}
blr_f1_pred <- adelie_chinstrap %>% 
  mutate(predict(blr_tidyfit_f1, .)) %>% 
  mutate(predict(blr_tidyfit_f1, ., type = 'prob'))

blr_f1_pred %>% 
  roc_curve(truth = species, .pred_Adelie) %>% 
  autoplot()

blr_f2_pred <- adelie_chinstrap %>% 
  mutate(predict(blr_tidyfit_f2, .)) %>% 
  mutate(predict(blr_tidyfit_f2, ., type = 'prob'))

blr_f2_pred %>% 
  roc_curve(truth = species, .pred_Adelie) %>% 
  autoplot()
```















